Software Citation Principles

### 2016-03-03 sc-principles/software-citation-principles.tex
some trimming of new text in 1 done, still need to add Mayernik cite and discuss

\katznote{add cites to James's study how what people cite software and Matthew Mayernik's similar one as part of motivation}

>>>>

\katznote{as part of motivation, James added cite to his study how what people cite software.  Dan to add cite of Matthew Mayernik's similar one, and perhaps trim text/table for space.}

Software and other digital resources currently appears in publications in very inconsistent ways.
For example, a random sample of 90 articles in the biology literature found seven different ways that software was mentioned, including simple names in the full-text, URLs in footnote, and different kinds of mentions in references lists: project names or websites, user manuals, publications that describe or introduce the software~\cite{howison2015jasist}. Table \ref{tab:mentions} shows examples of these varied forms of software mentions and the frequency with which they were encountered. 
%Even in the top journals (top 10 by impact factor), only 38\% of mentions included an entry of some sort in the reference list and fully 36\% were informal, following no recognizable style or format. 
Many of these kinds of mentions fail to perform the functions needed of citations, and their very diversity and frequent informality undermines the integration of software work into bibliometrics and other analyses. 
In some ways the situation is not too surprising: practices are diverse and informal because authors receive little guidance from editors, style guides, and their scientific communities, as well as weak support from citation software (such as Endnote and Zotero).
% In Howison and Bullard's study only 24\% of journals provided any guidance and software is treated inconsistently by style guides, if it appears at all.

\rowcolors{2}{white}{gray!25}
\begin{table}[htbp]
\caption{Varieties of software mentions in publications, from  \cite{howison2015jasist}} 
\centering
\scriptsize\setlength{\tabcolsep}{2.5pt}
\begin{tabular}{@{}l l l } % p{0.6\textwidth}@{}}
\toprule
Mention Type & \multicolumn{2}{c}{Count (n=286)} \\ % & Example \\
\midrule
Cite to publication & 105 & 37\% \\ %& \ldots was calculated using biosys (Swofford \& Selander 1981). \newline Ref: Swofford DL, Selander RB (1981) biosys-1: a Fortran program for the comprehensive analyses of electrophoretic data in population genetics and systematics. \emph{Journal of Heredity}, 72, 281–283.\\
Cite to users manual & 6 & 2\% \\ %& \ldots as analyzed by the BIAevaluation software (Biacore, 1997). \newline Ref: Biacore, I. (1997). BIAevaluation Software Handbook, version 3.0 (Uppsala, Sweden: Biacore, Inc) \\
Cite to name or website & 15 & 5\% \\ %& \ldots using the program Autodecay version 4.0.29 PPC (Eriksson 1998). \newline Ref: ERIKSSON, T. 1998. Autodecay, vers. 4.0.29 Stockholm:Department of Botany. \\
Instrument-like & 53 & 19\% \\ %& \ldots calculated by t-test using the Prism 3.0 software (GraphPad Software, San Diego, CA, USA). \\
URL in text & 13 & 5\% \\ %& \ldots freely available from http://www.cibiv.at/software/pda/. \\
In-text name only & 90 & 31\% \\ %& \ldots were analyzed using MapQTL (4.0) software. \\
Not even name & 4 & 1\% \\ %& \ldots was carried out using software implemented in the Java programming language. \\
\bottomrule
\end{tabular}
\label{tab:mentions}
\end{table}%
